Functory: Distributed Computing \\ for the Common Man\thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2010-10-21 doc/jfla-2011/main.tex
*** empty log message ***

\paragraph{Calcul distribué.}
Initialement inspirée par la bibliothèque
MapReduce\footnote{Ironiquement, l'approche de Google a été elle-même
  inspirée par la programmation fonctionnelle.} de
Google~\cite{mapreduce}, notre approche lui emprunte beaucoup de
vocabulaire.
La bibliothèque \functory\ est centrée autour de la notion de
\emph{tâche}. Les tâches représentent les calculs atomiques pouvant
être réalisés de manière indépendante.
Elles sont traitées par un \emph{patron} et des \emph{ouvriers}
(respectivement \emph{master} and \emph{workers} en anglais).
Les ouvriers représentent les capacités de calcul qui effectuent les
tâches. Ils sont
matérialisés par un ou plusieurs programmes, s'exécutant en parallèle
sur une ou plusieurs machines.
Le rôle du patron consiste à distribuer les tâches auprès des ouvriers
et à récolter les résultats. Il est matérialisé par un unique
programme s'exécutant de manière séquentielle.

@@

\hrulefill ICI\hrulefill

\medskip
These five functions are the most natural derivations of the
\of{compute} functions. There are obviously other possible derivations
(e.g. a variant of \of{map_fold} where only \of{fold} is meaningful);
in most cases, they could be derived in a straightforward way from our
generic API.

@@

% TODO cannot be instantiated in the monomorphic case

\section{Illustrative Case Studies}\label{sec:studies}

In this section, we show how to use \functory\ on several case studies.
We focus here on the use of the API. Experimental results of these
case studies are given in Section~\ref{sec:experiments}.
The source code for all these case studies is contained in the
distribution, in sub-directory \texttt{tests}.

\subsection{Matrix Multiplication}\label{sec:matrix}

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to p-1 do
      for k = 0 to m-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
assuming addition and multiplication over coefficients to be 
denoted $+$ and $\times$ respectively.
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to p-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

Using the sequential implementation provided by \of{Functory} is as
simple as including the following line in the code:
\begin{ocaml}
  open Sequential
\end{ocaml}
Then the computation is roughly similar to the standard multiplication
described above.

Now, let us assume we want to make use of a 4-core machine to perform
the same computation. 
This is achieved by just replacing the line above by
\begin{ocaml}
  open Cores
  let () = set_number_of_cores 4
\end{ocaml}
while the rest of the code remains unchanged.

Later, we may want to use a network of machines instead, for example
two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
cores respectively. Similar to the above, we turn the above lines into
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Same
\end{ocaml}
Here, \of{Same} is a module which is to be used when master and worker
are running the same executable. It provides a \of{compute} function
which has the same signature as in modules \of{Sequential} and
\of{Cores}, so that the rest of the code remains unchanged. Master and
worker processes are distinguished at run-time using an environment
variable \texttt{WORKER} which is set/unset.

If we need to write two different programs for master and worker, for
reason of binary incompatibility or any other reason, the library API
is providing functions to do that. If master and worker are still
compiled with the same version of \Ocaml, we use the \of{Poly} module
which provides a polymorphic API. Let us start with the worker program.
It now looks like:
\begin{ocaml}
  open Poly
  let worker (ai, bj) = ...
  let () = Worker.compute worker ()
\end{ocaml}
The \of{Worker.compute} function enters a loop which waits for
tasks sent by the master and returns results computed using \of{worker}.
The master program is almost the same as before. First, we replace
module \of{Same} with module \of{Poly}:
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Poly
\end{ocaml}
Tasks and \of{master} function are unchanged:
\begin{ocaml}
  let tasks = ...
  let master (_, (i,j)) r = ...
\end{ocaml}
Finally, we start the computation with \of{Master.compute}, which does
not have a \of{worker} parameter anymore:
\begin{ocaml}
  let () = Master.compute ~master tasks
\end{ocaml}

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string t = string_of_int (worker (task_of_string i))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

\subsection{N-queens}\label{sec:n-queens}

@@

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7  &  11.3\\
 (4 cores)     &   12.3  &   6.1\\
 (6 cores)     &    8.6  &   4.3\\
 (8 cores)     &    8.0  &   3.5\\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.

>>>>

\paragraph{Calcul distribué.}
Initialement inspirée par la bibliothèque
MapReduce\footnote{Ironiquement, l'approche de Google a été elle-même
  inspirée par la programmation fonctionnelle.} de
Google~\cite{mapreduce}, notre approche lui emprunte beaucoup de
vocabulaire.
La bibliothèque \functory\ est centrée autour de la notion de
\emph{tâche}. Les tâches représentent les calculs atomiques pouvant
être réalisés de manière indépendante.
Elles sont traitées par un \emph{patron} et des \emph{ouvriers}
(respectivement \emph{master} et \emph{workers} en anglais).
Les ouvriers représentent les capacités de calcul qui effectuent les
tâches. Ils sont
matérialisés par un ou plusieurs programmes, s'exécutant en parallèle
sur une ou plusieurs machines.
Le rôle du patron consiste à distribuer les tâches auprès des ouvriers
et à récolter les résultats. Il est matérialisé par un unique
programme s'exécutant de manière séquentielle.

@@

% \hrulefill ICI\hrulefill

%\medskip
Ces cinq fonctions se dérivent facilement de la fonction \of{compute}.
On peut évidemment en imaginer d'autres, comme par exemple une variante
où seule la fonction \of{fold} est significative.
L'utilisateur de \functory\ ne devrait pas avoir de mal à les dériver
lui-même de la fonction \of{compute} ou des fonctions du type
\of{map_fold} ci-dessus.

@@

\section{Études de cas}\label{sec:studies}

Cette section présente plusieurs études de cas que nous avons
réalisées avec \functory. On se concentre ici sur l'utilisation de la
bibliothèque ; les résultats expérimentaux sont présentés plus loin
(section~\ref{sec:experiments}). Le code source de tous les exemples
ci-dessus est contenu dans la distribution de \functory, dans le
sous-répertoire \texttt{tests}.

\subsection{Multiplication de matrices}\label{sec:matrix}

Comme premier exemple, on considère la multiplication de deux matrices
\of{a} et \of{b}, de tailles respectives
$\of{n}\times\of{p}$ et $\of{p}\times\of{m}$. 
Le résultat sera stocké dans une matrice
\of{c} de taille $\of{n}\times\of{m}$.
On suppose que \of{a}, \of{b} et \of{c} sont contenues dans des
variables globales.
En supposant en outre \of{a} organisée en lignes et \of{b} en colonnes, une
multiplication séquentielle s'écrirait ainsi 
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to m-1 do
      for k = 0 to p-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
l'addition et la multiplication des coefficients étant notées
respectivement $+$ et $\times$.
La complexité de ce code est clairement $O(\of{n}\times\of{m}\times\of{p})$,

Une manière évidente de distribuer ce calcul consiste à faire de la
boucle interne sur \of{k} une tâche élémentaire.
On construit alors une liste de $\of{n} \times \of{m}$ tâches, de la
manière suivante :
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to m-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Chaque tâche est une paire formée d'une part d'une ligne \of{a.(i)} et
d'une colonne \of{b.(j)}, et d'autre part d'une position \of{(i,j)}.
L'ouvrier reçoit la première composante et calcule le produit scalaire.
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
La patron est une fonction d'une ligne, qui reçoit le résultat \of{r}
renvoyé par l'ouvrier et remplit la matrice \of{c}, en fonction de la
position contenue dans la seconde composante de la tâche. Aucune 
nouvelle tâche n'est produite.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Au final, l'ensemble du calcul est lancé par un appel à la fonction
\of{compute}.
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}

Utiliser la version séquentielle de \functory\ est aussi simple que
d'ajouter \of{open Sequential} au début du code. Le calcul est alors
similaire à la multiplication usuelle donnée plus haut. Cela peut être
néanmoins utile pour vérifier la correction du code avant de chercher
à distribuer le calcul.

Supposons maintenant que l'on veuille utiliser une machine 4 \coeurs\
pour effectuer le calcul. Il suffit alors de remplacer la ligne
\of{open Sequential} par les deux lignes suivantes.
\begin{ocaml}
  open Cores
  let () = set_number_of_cores 4
\end{ocaml}
Le reste du code est inchangé.

Supposons enfin que l'on souhaite utiliser plutôt des machines
présentes sur le réseau, par exemple deux machines appelées
\texttt{orcus} et \texttt{belzebuth} offrant respectivement 4 et 8 \coeurs.
Il suffit de remplacer les deux lignes de code ci-dessus par les suivantes.
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Same
\end{ocaml}
On utilise ici le sous-module \of{Same} de \of{Network}, qui permet
d'utiliser le même exécutable pour le patron et les ouvriers. Ce
module fournit toujours une fonction \of{compute} de même signature
que dans les modules \of{Sequential} et \of{Cores}, et le reste du
code est toujours inchangé. Le patron et les ouvriers sont distingués
à l'exécution par la présence de la variable d'environnement
\texttt{WORKER}.

Si on a besoin d'écrire deux programmes différents pour le patron et
les ouvriers, pour des raisons d'incompatibilité de binaires ou tout
autre raison, l'interface de \functory\ permet de le faire.
Si les deux programmes sont compilés avec la même version d'\Ocaml, on
utilise le module \of{Poly}.
Commençons par l'ouvrier. Son code prend la forme suivante.
\begin{ocaml}
  open Network.Poly
  let worker (ai, bj) = ...
  let () = Worker.compute worker ()
\end{ocaml}
La fonction \of{Worker.compute} entre dans une boucle qui attend les
tâches envoyées par le patron et renvoie les résultats calculés par
la fonction \of{worker}. Le code du patron, quant à lui, est quasiment
le même qu'auparavant. On commence par remplacer \of{Same} par
\of{Poly}.
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Poly
\end{ocaml}
La construction des tâches et la fonction \of{master} sont inchangées.
\begin{ocaml}
  let tasks = ...
  let master (_, (i,j)) r = ...
\end{ocaml}
Enfin on lance le calcul avec la fonction \of{Master.compute}, qui ne
prend plus de fonction \of{worker} en argument.
\begin{ocaml}
  let () = Master.compute ~master tasks
\end{ocaml}

Quand le patron et les ouvriers sont compilés avec des versions
différentes d'\Ocaml, la bibliothèque fournit une interface
ne permettant plus que l'échange de chaînes de caractères.
En conséquence, il faut convertir les tâches et les résultats depuis
et vers des chaînes, dans les deux programmes.
L'ouvrier modifié prend alors la forme suivante.
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string s = string_of_coeff (worker (task_of_string s))
  let () = Worker.compute worker_string ()
\end{ocaml}
Le patron est modifié de la même façon.
The master program is modified in a similar way.
On remplace \of{Poly} par \of{Mono} et on encode/décode les tâches et
résultats.
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
Les quatre fonctions de conversion \of{string_of_}\{\of{task,coeff}\} et
\{\of{task,coeff}\}\of{_of_string} sont à la charge de l'utilisateur.

\subsection{Autres études de cas}

Nous présentons ici trois autres études de cas, plus rapidement.

\paragraph{N-reines.}\label{sec:n-queens}

@@

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 \phantom{(0.00)} &  20.2 \phantom{(0.00)} \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
 (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
 (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
 (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.

### 2010-10-21 doc/report/main.tex
*** empty log message ***

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to p-1 do
      for k = 0 to m-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
assuming addition and multiplication over coefficients to be 
denoted $+$ and $\times$ respectively.
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to p-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to m-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

@@

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string t = string_of_int (worker (task_of_string i))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

@@

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7  &  11.3\\
 (4 cores)     &   12.3  &   6.1\\
 (6 cores)     &    8.6  &   4.3\\
 (8 cores)     &    8.0  &   3.5\\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.

>>>>

As a first example, let us consider matrix multiplication.
We assume two matrices \of{a} and \of{b}, respectively of size
$\of{n}\times\of{p}$ and $\of{p}\times\of{m}$, given as input, as well as
a matrix \of{c} of size $\of{n}\times\of{m}$ to store the result.
Assuming \of{a} being a row-matrix and \of{b} a column-matrix,
the standard matrix multiplication would be as follows:
\begin{ocaml}
  for i = 0 to n-1 do
    for j = 0 to m-1 do
      for k = 0 to p-1 do
	c.(i).(j) <- c.(i).(j) + a.(i).(k) * b.(j).(k)
      done
    done
  done  
\end{ocaml}
and runs in $O(\of{n}\times\of{p}\times\of{m})$, 
assuming addition and multiplication over coefficients to be 
denoted $+$ and $\times$ respectively.
An easy way to distribute this computation is obviously to turn each
computation of the inner loop over \of{k} into a task.
To do so, we first prepare the list of tasks:
\begin{ocaml}
  let tasks = 
    let l = ref [] in
    for i = 0 to n-1 do for j = 0 to m-1 do
      tasks := ((a.(i), b.(j)), (i,j)) :: !tasks
    done done;
    !l
\end{ocaml}
Each task consists of row \of{a.(i)} and column \of{b.(j)} as first
tuple, together with position \of{(i,j)} as a second tuple.
The worker is receiving the first tuple as argument and computes
the dot product:
\begin{ocaml}
  let worker (ai, bj) =
    let c = ref 0 in
    for k = 0 to p-1 do c := !c + ai.(k) * bj.(k) done;
    !c
\end{ocaml}
The master is a one line function which receives the result \of{r}
from the worker and simply stores it into \of{c}, according to the
position contained in the input task. It produces no new task.
\begin{ocaml}
  let master (_, (i,j)) r = c.(i).(j) <- r; []
\end{ocaml}
Finally, the overall computation is started by invoking \of{compute}
as follows:
\begin{ocaml}
  let () = compute ~worker ~master !tasks
\end{ocaml}
The total number of tasks is $\of{n}\times\of{m}$.

@@

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
tasks and results to and from strings in both master and worker.
The modified worker program then looks as follows:
\begin{ocaml}
  open Mono
  let worker (ai, bj) = ...
  let worker_string s = string_of_coeff (worker (task_of_string s))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode coefficients
as strings, as follows:
\begin{ocaml}
  let tasks = ... string_of_task ...
  let master (_, (i,j)) r = c.(i).(j) <- coeff_of_string r; []
\end{ocaml}
where \of{string_of_task} and \of{task_of_string} are user-defined
functions to convert tasks to and from strings.

@@

We compare the performances of two different implementations. In the
first one, called \of{mm1}, each task consists of the computation of a single
coefficient of the resultant matrix, as described in Section~\ref{sec:matrix}.
In the second one, called \of{mm2}, each task consists of the
computation of a whole row of the resultant matrix.
As a consequence, the total number of tasks in
$\of{n}\times\of{m}=10,000$ for \of{mm1} and only $\of{n}=100$ for \of{mm2}.
The experimental results (in seconds) are tabulated below.
\begin{center}
  \begin{tabular}{|r|r|r|}
    \hline
    & \of{mm1}       & \of{mm2}  \\
    & (10,000 tasks) & (100 tasks) \\
    \hline\hline
\of{Sequential} & 20.3 &  20.2 \\
\hline
 \of{Cores} 
 (2 cores)     &   22.7 (0.89) &  11.3 (1.79) \\
 (4 cores)     &   12.3 (1.65) &   6.1 (3.31) \\
 (6 cores)     &    8.6 (2.36) &   4.3 (4.70) \\
 (8 cores)     &    8.0 (2.54) &   3.5 (\underline{5.77}) \\
 \hline
  \end{tabular}
\end{center}
We do not include results for the network configuration, as they do
not achieve any benefit with respect to the sequential
implementation. The reason is that the communication cost dominates the
computation cost in such a way that the total execution time is
always greater than 30 seconds. Indeed, irrespective of the
implementation (\of{mm1} or \of{mm2}), the total size of the
transmitted data is 
$O(\of{n}\times\of{m}\times\of{p})$, which in our case amounts to
billions of bytes.
